---
title: "Midterm 1 W24"
author: "Maya Fujita"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code must be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above. 

Your code must knit in order to be considered. If you are stuck and cannot answer a question, then comment out your code and knit the document. You may use your notes, labs, and homework to help you complete this exam. Do not use any other resources- including AI assistance.  

Don't forget to answer any questions that are asked in the prompt!  

Be sure to push your completed midterm to your repository. This exam is worth 30 points.  

## Background
In the data folder, you will find data related to a study on wolf mortality collected by the National Park Service. You should start by reading the `README_NPSwolfdata.pdf` file. This will provide an abstract of the study and an explanation of variables.  

The data are from: Cassidy, Kira et al. (2022). Gray wolf packs and human-caused wolf mortality. [Dryad](https://doi.org/10.5061/dryad.mkkwh713f). 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the wolves data
In these data, the authors used `NULL` to represent missing values. I am correcting this for you below and using `janitor` to clean the column names.
```{r message=FALSE, warning=FALSE}
wolves <- read.csv("data/NPS_wolfmortalitydata.csv", na = c("NULL")) %>% clean_names()
```

## Questions
Problem 1. (1 point) Let's start with some data exploration. What are the variable (column) names?  
```{r}
names(wolves)
```


Problem 2. (1 point) Use the function of your choice to summarize the data and get an idea of its structure.  
```{r}
glimpse(wolves)
```

Problem 3. (3 points) Which parks/ reserves are represented in the data? Don't just use the abstract, pull this information from the data.  
```{r}
wolves$park <- as.factor(wolves$park)
```
```{r}
wolves %>% 
  count(park)
```
The five parks that are represented in the data are DENA, GNTP, VNP, YNP, YUCH

Problem 4. (4 points) Which park has the largest number of wolf packs?

```{r}
wolves %>% 
  group_by(park) %>% 
  summarize(n_packs = n())
```
DENA park had the highest number of wolf packs with a total of 340.

Problem 5. (4 points) Which park has the highest total number of human-caused mortalities `mort_all`?
```{r}
wolves %>% 
  group_by(park) %>% 
  summarize(max_mortalities=max(mort_all)) %>% 
  arrange(desc(max_mortalities))
```
YUCH had the highest total number of human-caused mortalities.

The wolves in [Yellowstone National Park](https://www.nps.gov/yell/learn/nature/wolf-restoration.htm) are an incredible conservation success story. Let's focus our attention on this park.  

Problem 6. (2 points) Create a new object "ynp" that only includes the data from Yellowstone National Park.
```{r}
ynp <- wolves %>% 
  filter(park == "YNP")
```


Problem 7. (3 points) Among the Yellowstone wolf packs, the [Druid Peak Pack](https://www.pbs.org/wnet/nature/in-the-valley-of-the-wolves-the-druid-wolf-pack-story/209/) is one of most famous. What was the average pack size of this pack for the years represented in the data?
```{r}
ynp$biolyr <- as.factor(ynp$biolyr)
ynp %>% 
  filter(pack == "druid") %>% 
  group_by(biolyr) %>% 
  summarize(avg_packsize = mean(packsize_aug))
```

Problem 8. (4 points) Pack dynamics can be hard to predict- even for strong packs like the Druid Peak pack. At which year did the Druid Peak pack have the largest pack size? What do you think happened in 2010?
```{r}
ynp %>% 
  filter(pack == "druid") %>% 
  group_by(biolyr) %>% 
  summarize(max_packsize = max(packsize_aug)) %>% 
  arrange(desc(max_packsize))
```
In the year 2001, the Druid Peak pack had the largest pack size o 37 when measured in August. 
```{r}
ynp %>% 
  filter(pack == "druid") %>% 
  filter(biolyr == "2008" |biolyr == "2009"|biolyr == "2010")
```
With a closer look at the data from 2010, and two years before, the pack size dramatically decressed in two years, going from 21 to 0. However, there was no indication of an increate in mortalities due to human causes (mort_all), as for all three years they were 0. So this leads me to believe there was a natural reason for the decline in this pack. With a google search, I found that in 2009 to 2010, the pack contracted mange, which is a disease wolves can get causing many of the wolves to die from starvation and skirmishes, which would explain the pack size for 2010. (https://yollowstonewolves.fandom.com/wiki/Druids_Pack)

Problem 9. (5 points) Among the YNP wolf packs, which one has had the highest overall persistence `persisty1` for the years represented in the data? Look this pack up online and tell me what is unique about its behavior- specifically, what prey animals does this pack specialize on?  
```{r}
ynp %>% 
  group_by(pack) %>% 
  filter(!is.na(persisty1)) %>% 
  summarise(total_persistance=sum(persisty1)) %>% 
  arrange(desc(total_persistance))
```
The mollies pack had the largest total persistence over the years represented in the data. According the my google search, this pack learned to specialize in hunting bison, making them relatively successful compared to other packs. (https://www.yellowstonewolf.org/yellowstones_wolves.php?pack_id=6)

Problem 10. (3 points) Perform one analysis or exploration of your choice on the `wolves` data. Your answer needs to include at least two lines of code and not be a summary function. 

I was interested to know for all the wolf packs in DENA park, which year saw the largest average pack sizes across all packs.
```{r}
wolves$biolyr <- as.factor(wolves$biolyr)
  
```

```{r}
wolves %>% 
  group_by(biolyr) %>% 
  filter(park == "DENA") %>% 
  filter(!is.na(packsize_aug)) %>%
  summarize(mean_packsize=mean(packsize_aug)) %>% 
  arrange(desc(mean_packsize))
```
I found that in 1990, the average pack sizes across all packs in DENA park were the largest with an average of 13.
