---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI refers to the simulation of human intelligence in machines that are programmed to think and learn like humans. It encompasses a wide range of technologies and techniques, including machine learning, natural language processing, computer vision, and more. The goal of AI is to create systems that can perform tasks that typically require human intelligence, such as problem-solving, speech recognition, and decision-making.
2. What is ChatGPT?
ChatGPT is a language model developed by OpenAI, based on the GPT (Generative Pre-trained Transformer) architecture. It is designed to generate human-like text and engage in natural language conversations. The model is trained on diverse internet text data and can be used for various natural language processing tasks, including chat-based applications, content generation, and language understanding.
3. What is Github Copilot?
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It is designed to assist developers by providing code suggestions and auto-completions as they write code. GitHub Copilot is powered by OpenAI's Codex, a language model that has been trained on a vast amount of code from publicly available repositories. It aims to improve productivity by helping developers write code faster and reducing the cognitive load associated with coding tasks.

4. What is tidyverse?
Tidyverse: The tidyverse is a collection of R packages for data science and statistical computing that share a common philosophy and syntax. It is designed to make data analysis and visualization in R more efficient, consistent, and user-friendly. The tidyverse packages are built around the idea of tidy data, which is a structured and standardized way of organizing and manipulating data.

Some key packages within the tidyverse include:

dplyr: Provides a set of functions for data manipulation and transformation.
ggplot2: A powerful and flexible package for creating static, animated, and interactive data visualizations.
tidyr: Focuses on data tidying and reshaping, helping to clean and structure data for analysis.
readr: Offers functions for reading and writing data in various formats, improving data import/export.
purrr: A functional programming toolkit that enhances the way data manipulation functions are used.
tibble: Introduces the tibble data frame, an enhanced version of the traditional data frame in R.
The tidyverse is popular among data scientists and analysts working in R, as it promotes a consistent and expressive workflow for data analysis and visualization tasks.





## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("data/netflix_titles.csv")

```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}


```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- filter(netflix, type == "Movie")

```


Which director has the most movies on netflix?
```{r}
netflix_movies %>%
  group_by(director) %>% 
  count(director) %>% 
  arrange(-n)
  


```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*
seperate each director into own row
```{r}

```



Which director had the most movies by release year and remove the NA's?
```{r}


```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix remove the nas
```{r}
netflix %>%
  na.omit() %>%
  group_by(director) %>%
  summarise(number_of_movies = n()) %>%
  arrange(desc(number_of_movies)) %>%
  head(10) %>%  # Adjust the number to show more or fewer top directors
  ggplot(aes(x = reorder(director, -number_of_movies), y = number_of_movies)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(
    title = "Top Directors with the Most Movies on Netflix (after removing NAs)",
    x = "Director",
    y = "Number of Movies"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*
```{r}
ui <- fluidPage(
  titlePanel("Netflix Movie Analysis"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("release_year", "Select Release Year", c(2008, 2009, 2010))
    ),
    
    mainPanel(
      plotOutput("bar_plot")
    )
  )
)

# Server logic
server <- function(input, output) {
  
  output$bar_plot <- renderPlot({
    # Filter data based on selected release year
    filtered_data <- netflix %>%
      na.omit() %>%
      filter(year(as.Date(release_date)) == input$release_year) %>%
      group_by(director) %>%
      summarise(number_of_movies = n()) %>%
      arrange(desc(number_of_movies)) %>%
      head(10)
    
    # Create bar plot
    ggplot(filtered_data, aes(x = reorder(director, -number_of_movies), y = number_of_movies)) +
      geom_bar(stat = "identity", fill = "skyblue") +
      theme_minimal() +
      labs(
        title = paste("Top Directors with the Most Movies on Netflix (", input$release_year, ")"),
        x = "Director",
        y = "Number of Movies"
      ) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  })
}

# Run the Shiny app
shinyApp(ui, server)
```





## That's it! Let's take a break and then move on to working on group projects.
